---
title: "PUBH 691 Summarizing Data"
author: "Alexandria Albers"
date: "2022-10-04 and OCt 6,2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rio)
library(janitor)
library(skimr)
library(stargazer)
library(knitr)
```

```{r}
# Read in the NHANES data files
nhanes_raw <- import("input/nhanes_raw.csv") %>% 
   clean_names() %>% 
   mutate(across(everything()), na_if(., "")) %>%
   mutate(across(where(is.character), as.factor))

```

```{r}
#install.packages("skimr")

skim(nhanes_raw)
```

```{r}
skim(nhanes_raw$health_gen)

summary(nhanes_raw$health_gen)[[2]] #putting 2 sets of brackets and number pulls an individual level of a factor

summary(nhanes_raw$health_gen)[["Fair"]]

skim(nhanes_raw$health_gen)[["top_counts"]]

skim_output <-skim(nhanes_raw$health_gen)
summary_output <- summary(nhanes_raw$health_gen)

skim(nhanes_raw$health_gen)[["factor.top_counts"]]
```

```{r}
table(nhanes_raw$marital_status, nhanes_raw$health_gen)

table(nhanes_raw$marital_status, nhanes_raw$health_gen, useNA = "always")

ftable(nhanes_raw$marital_status, nhanes_raw$health_gen, nhanes_raw$gender)

#ftable is flat table - gives you nicer output for 3+ variables

my_table <- table(nhanes_raw$marital_status, nhanes_raw$health_gen)

my_table

margin.table(my_table, 1) #comma 1 gives you freq across table for 1st variable
margin.table(my_table, 2) #will give you for 2nd variable

prop.table(my_table) #proportions/percentages within the table
prop.table(table(nhanes_raw$marital_status, nhanes_raw$health_gen))

prop.table(my_table, 1) #each row now adds to 100%
prop.table(my_table, 2) #each column adds up to 10%
```

```{r}
nhanes_raw %>% 
  tabyl(education, show_na=F) %>% 
  adorn_pct_formatting()

nhanes_raw %>% 
  tabyl(education, health_gen, show_na=F) %>% 
  adorn_totals(where = "row") #adding total at the end, but getting column totals

nhanes_raw %>% 
  tabyl(education, health_gen, show_na=F) %>% 
  adorn_totals(where = "col") #getting row totals by adding a col total

nhanes_raw %>% 
  tabyl(education, health_gen, show_na=F) %>% 
  adorn_totals(where = "both")

nhanes_raw %>% 
  tabyl(education, health_gen, show_na=F) %>% 
  adorn_totals(where = "both") %>% 
  adorn_percentages(denominator = "row") #need to specify how you want den calculated
                                              #for "row", each row adds up to 100%
                                            #which needs to be specified if 2 diff vars

nhanes_raw %>% 
  tabyl(education, health_gen, show_na=F) %>% 
  adorn_totals(where = "both") %>% 
  adorn_percentages(denominator = "row") %>% 
  adorn_pct_formatting()

#if you want numbers and percentages
nhanes_raw %>% 
  tabyl(education, health_gen, show_na=F) %>% 
  adorn_totals(where = "both") %>% 
  adorn_percentages(denominator = "row") %>% 
  adorn_pct_formatting() %>% 
  adorn_ns(position = "front")

#if you want to label and look nicer
nhanes_raw %>% 
  tabyl(education, health_gen, show_na=F) %>% 
  adorn_totals(where = "both") %>% 
  adorn_percentages(denominator = "row") %>% 
  adorn_pct_formatting() %>% 
  adorn_ns(position = "front") %>% 
  adorn_title(row_name = "Education",
              col_name = "General Health")
```


##flex table package
##gt summary <- try this one first!
## look at epi handbook
##stargazer

#from patrick...
```{r}
my_tabyl <- nhanes_raw %>% 
  tabyl(education, health_gen, show_na = F) %>% 
  adorn_totals(where ="both") %>% 
  adorn_percentages(denominator = "row") %>% 
  adorn_pct_formatting() %>% 
  adorn_ns(position = "front") %>% 
  adorn_title(row_name = "Education",
              col_name = "General Health")

library(stargazer)
stargazer(my_tabyl, type = "html", out = "tabyl.html", summary = F)

```

```{r}
nhanes_raw %>% 
  summarize("N" = n(),
            "Mean Age" = round(mean(age), digits=2),
            "SD Age" = round(sd(age), digits=2))

age_summary <- nhanes_raw %>% 
  filter(!is.na(diabetes)) %>% 
  group_by(diabetes) %>% 
  summarize("n" = n(),
            "Mean Age" = round(mean(age), digits=2),
            "SD Age" = round(sd(age), digits=2),
            "Min Age" = min(age),
            "25% Age" = quantile(age, 0.25),
            "75% Age" = quantile(age, 0.75),
            "Median Age" = median(age),
            "Max Age" = max(age)) %>% 
  rename("Diabetes Status" = diabetes)

age_summary

#if you have specific summary stats you want, the summarize fn is the
#best way to do this
```


```{r}
age_summary_long <-age_summary %>%   #from wide dataset to longer dataset
  pivot_longer(cols = n:'Max Age',
               names_to = "Statistic",
               values_to = "Result")

age_summary_long
```

```{r}
age_summary_wide <- age_summary_long %>% 
  pivot_wider(names_from = "Statistic",
              values_from = "Result")

age_summary_wide
```

#combining columns
```{r}
age_summary_long2 <- age_summary_long %>% 
  mutate(diabetes = "Diabetes") %>% 
  unite("Health Status", diabetes, 'Diabetes Status', sep = " ") #the default is to get rid of columns that were united,
                                                                #but could put remove=F would keep the cols in the dataset

age_summary_long2
```


#sep col into 2
```{r}
age_summary_long3 <- age_summary_long2 %>% 
  separate(col = "Health Status", into = c("Diabetes", "Diabetes Status"), sep = " ", 
           remove = F)

age_summary_long3
```


#Conditional Statistics
```{r}
bp_dia_gen_summary <- nhanes_raw %>% 
  group_by(gender) %>% 
  summarize("Mean SBP Diabetes" = mean(bp_sys_ave[diabetes == "Yes"], na.rm=T),   # so we do want mean SBP, but ONLY when diabetes=Y
            "Mean SBP No Diabetes" = mean(bp_sys_ave[diabetes == "No"], na.rm=T),
            "Mean DBP Diabetes" = mean(bp_dia_ave[diabetes =="Yes"], na.rm=T),
            "Mean DBP No Diabetes" = mean(bp_dia_ave[diabetes =="No"], na.rm=T)) %>% 
  rename(Gender=gender)

bp_dia_gen_summary
kable(bp_dia_gen_summary) #in knitr library
```


