---
title: "Final Figures for MS"
output: html_document
date: "2022-08-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
install.packages("haven")
install.packages("sas7bdat")
install.packages("writexl")
install.packages ("fmtr")
install.packages("stringr")
install.packages ("forcats")
```

```{r}
library (haven)
library (sas7bdat)
library (readxl)
library (writexl)
library (tidyr)
library (ggplot2)
library (dplyr)
library (magrittr)
library (fmtr)
library (stringr)
library (forcats)
library (ggplot2)
library (common)
library (readr)
```



#import sas dataset into r
```{r}
df<- read_sas("redcap.sas7bdat")
```

#read excels in for 3 figures
#ECH barriers, ECH parent concerns, CV19 barriers

```{r}
ECHbarriers <- read_excel ("echbarriers.xlsx")
ECHconcerns <- read_excel ("echparentconcerns.xlsx")
CV19barriers <- read_excel ("cv19barriers.xlsx")
ECHbarriers_ruca <- read_excel ("echbarriers_ruca.xlsx")
ECHconcerns_ruca <- read_excel ("echparentconcerns_ruca.xlsx")
CV19barriers_ruca <- read_excel ("cv19barriers_ruca.xlsx")
```


```{r}
write_xlsx(ECHbarriers, "C:/Users/alber/OneDrive/Desktop/R/echbarriers.xlsx")
```


##Plot for ALL providers (not broken out by RUCA categories)
##Early childhood barriers
```{r, include=FALSE}
ff <- factor (ECHbarriers$Barrier, levels = c("Not a barrier (early childhood)", "Barrier (early childhood)"))

Statements <- factor (ECHbarriers$Options, levels = c( 
                                  "Inability of our facility to schedule visits more than a month or two out", 
                                          "Not having enough nursing or other support staff in our facility",
                                   "Lack of automated prompts in our EHR telling us which vaccines are due or overdue for the patient",
                                  "Transportation obstacles experienced by parents or families to get to our facility",
                                         "Parents not knowing when to bring their child in for routine well care",
                                
                 
                  "Parents choosing to delay or refuse some or all vaccines for their children"))
                                                                      

Fig2 <- ggplot(ECHbarriers, aes(x = Statements, y = Percent, fill = ff)) +
         geom_bar(stat = "identity", alpha = 1, width = 0.5) +
          coord_flip() +
          scale_y_continuous(name="Percentage of Providers", labels = function (y) paste0(y, "%")) +
    theme_minimal () +
    scale_x_discrete(labels = function (Statements) str_wrap (Statements, width = 25)) +
   theme(axis.text.x = element_text(face="bold", colour = "black",
                           size=14, angle=0),
          axis.text.y = element_text(face="bold", colour = "black",
                           size=11, angle=0)) +
  theme(legend.title=element_blank(),
        legend.text = element_text (size = 10, face="bold"),
        axis.title.y = element_blank (),
        axis.title.x = element_text (size = 16, face = "bold")) +
  theme (legend.position = "bottom") +
  theme (axis.title.x = element_text (margin = margin (t=20)))+
  theme (panel.grid.major = element_line (size = 0.5, linetype = "solid", color="grey"))+
 scale_fill_brewer(palette = "Paired") +
   guides (fill = guide_legend (reverse = TRUE))


#to create pdf, but need to name ggplot plot1, so first, you would
#do the plot1 <- ggplot.... , can also add a family = "Times" option
#in the first line, ex. pdf ("plot1.pdf", family = "Times")

pdf ("Fig2.pdf")
Fig2
dev.off ()
```



##Plot for ALL providers (not broken out by RUCA categories)
##Early childhood -provider reported parental concerns
```{r, include=FALSE}
ff <- factor (ECHconcerns$Amount, levels = c("Not a parental concern", "Parental concern"))

Statements <- factor (ECHconcerns$Options, levels = c( 
                                  "Concern that vaccines will weaken their child's immune system",
                 "Belief that vaccines are given when children are too young",
                 "Belief that vaccine-preventable diseases are not severe enough to warrant vaccination",
                 "Belief that their child is not at risk for vaccine-preventable diseases",
                 "Concern that their child will suffer immediate, short-term side effects, including pain",
                "General worries about vaccines without a specific concern",
                 "Concern that their child will suffer long-term complications from vaccines",
                 "Concern that children receive too many vaccines"))
                                                                      

Fig3 <- ggplot(ECHconcerns, aes(x = Statements, y = Percent, fill = ff)) +
         geom_bar(stat = "identity", alpha = 1, width = 0.5) +
          coord_flip() +
          scale_y_continuous(name="Percentage of Providers", labels = function (y) paste0(y, "%")) +
    theme_minimal () +
    scale_x_discrete(labels = function (Statements) str_wrap (Statements, width = 35)) +
   theme(axis.text.x = element_text(face="bold", colour = "black",
                           size=14, angle=0),
          axis.text.y = element_text(face="bold", colour = "black",
                           size=11, angle=0)) +
  theme(legend.title=element_blank(),
        legend.text = element_text (size = 10, face="bold"),
        axis.title.y = element_blank (),
        axis.title.x = element_text (size = 16, face = "bold")) +
  theme (legend.position = "bottom") +
  theme (axis.title.x = element_text (margin = margin (t=20)))+
  theme (panel.grid.major = element_line (size = 0.5, linetype = "solid", color="grey"))+
 scale_fill_brewer(palette = "Paired") +
   guides (fill = guide_legend (reverse = TRUE))


#to create pdf, but need to name ggplot plot1, so first, you would
#do the plot1 <- ggplot.... , can also add a family = "Times" option
#in the first line, ex. pdf ("plot1.pdf", family = "Times")

pdf ("Fig3.pdf")
Fig3
dev.off ()
```


##Plot for ALL providers (not broken out by RUCA categories)
##COVID-19 barriers
```{r, include=FALSE}
ff <- factor (CV19barriers$Barrier, levels = c("Not a barrier (COVID-19)", "Barrier (COVID-19)"))

Statements <- factor (CV19barriers$Options, levels = c( 
  "Transportation obstacles to getting a COVID-19 vaccine",
"Parents' or patients' fears that getting the COVID-19 vaccine will cause short-term side effects/allergic reaction",
"Parents' or patients' beliefs that the COVID-19 vaccine is being promoted for commercial gains of pharmaceutical companies",
"Parents' or patients' fears that getting the COVID-19 vaccine will lead to long-term health issues",
"Parents' or patients' concerns that the COVID-19 vaccines were rapidly developed and authorized"))
                                                                      

Fig4 <- ggplot(CV19barriers, aes(x = Statements, y = Percent, fill = ff)) +
         geom_bar(stat = "identity", alpha = 1, width = 0.5) +
          coord_flip() +
          scale_y_continuous(name="Percentage of Providers", labels = function (y) paste0(y, "%")) +
    theme_minimal () +
    scale_x_discrete(labels = function (Statements) str_wrap (Statements, width = 35)) +
   theme(axis.text.x = element_text(face="bold", colour = "black",
                           size=14, angle=0),
          axis.text.y = element_text(face="bold", colour = "black",
                           size=11, angle=0)) +
  theme(legend.title=element_blank(),
        legend.text = element_text (size = 10, face="bold"),
        axis.title.y = element_blank (),
        axis.title.x = element_text (size = 16, face = "bold")) +
  theme (legend.position = "bottom") +
  theme (axis.title.x = element_text (margin = margin (t=20)))+
  theme (panel.grid.major = element_line (size = 0.5, linetype = "solid", color="grey"))+
 scale_fill_brewer(palette = "Paired") +
   guides (fill = guide_legend (reverse = TRUE))


#to create pdf, but need to name ggplot plot1, so first, you would
#do the plot1 <- ggplot.... , can also add a family = "Times" option
#in the first line, ex. pdf ("plot1.pdf", family = "Times")

pdf ("Fig4.pdf")
Fig4
dev.off ()
```






































## Plot for early childhood barriers, broken out by RUCA categories (metro and nonmetro)
```{r, include=FALSE}
ff <- factor (ECHbarriers_ruca$Barrier, levels = c("Not a barrier (early childhood)", "Barrier (early childhood)"))

Fig2ruca_ech_barriers <- ggplot (ECHbarriers_ruca, aes (x=RUCA, y=Percent, fill= ff)) + 
  geom_bar (position = "stack", stat = "identity", width = 0.8) +
  facet_wrap (~fct_relevel (Rank, "Rank6", "Rank4", "Rank5", "Rank3", "Rank1", "Rank2"), nrow=6) +
  

  coord_flip() +

          scale_y_continuous(name="Percentage of Providers", labels = function (y) paste0(y, "%")) +
    theme_minimal () +
   theme(axis.text.x = element_text(face="bold", colour = "black",
                           size=14, angle=0),
          axis.text.y = element_text(face="bold", colour = "black",
                           size=13, angle=0)) +
  #geom_text (aes (label=paste0 (round (Percent, 1), "%")), size = 5, hjust = 0.8) +
  theme(legend.title=element_blank(),
        legend.text = element_text (size = 12, face="bold"),
        axis.title.y = element_blank (),
        axis.title.x = element_text (size = 16, face = "bold")) +
  theme (legend.position = "bottom") +
  theme (axis.title.x = element_text (margin = margin (t=20)))+
  theme (panel.grid.major = element_line (size = 0.5, linetype = "solid", color="grey"))+
  theme (panel.spacing = unit (2,"lines")) +
  theme (strip.text.x = element_blank () ) +
 scale_fill_brewer(palette = "Paired") +
   guides (fill = guide_legend (reverse = TRUE))


pdf ("Fig2ruca_ech_barriers.pdf")
Fig2ruca_ech_barriers
dev.off ()
```



##Plot for early childhood parental concerns question by ruca catego of metro and nonmetro
```{r, include=FALSE}
ff <- factor (ECHconcerns_ruca$Amount, levels = c("Not a parental concern", "Parental concern"))


Fig3ruca_ech_concerns <- ggplot (ECHconcerns_ruca, aes (x=RUCA, y=Percent, fill= ff)) + 
  geom_bar (position = "stack", stat = "identity", width = 0.8) +
  facet_wrap (~fct_relevel (Rank, "Rank6", "Rank5", "Rank8", "Rank1", "Rank3", "Rank2", "Rank7",
                            "Rank4"), nrow=8) +
  

  coord_flip() +

          scale_y_continuous(name="Percentage of Providers", labels = function (y) paste0(y, "%")) +
    theme_minimal () +
   theme(axis.text.x = element_text(face="bold", colour = "black",
                           size=14, angle=0),
          axis.text.y = element_text(face="bold", colour = "black",
                           size=13, angle=0)) +
  #geom_text (aes (label=paste0 (round (Percent, 1), "%")), size = 5, hjust = 0.8) +
  theme(legend.title=element_blank(),
        legend.text = element_text (size = 12, face="bold"),
        axis.title.y = element_blank (),
        axis.title.x = element_text (size = 16, face = "bold")) +
  theme (legend.position = "bottom") +
  theme (axis.title.x = element_text (margin = margin (t=20)))+
  theme (panel.grid.major = element_line (size = 0.5, linetype = "solid", color="grey"))+
  theme (panel.spacing = unit (2,"lines")) +
  theme (strip.text.x = element_blank () ) +
 scale_fill_brewer(palette = "Paired") +
   guides (fill = guide_legend (reverse = TRUE)) +
 


pdf ("Fig3ruca_ech_concerns.pdf")
Fig3ruca_ech_concerns
dev.off ()
```


#plot for COVID-19 barriers by RUCA categ of metro and nonmetro
```{r, include= FALSE}
ff <- factor (CV19barriers_ruca$Barrier, levels = c("Not a barrier (COVID-19)", "Barrier (COVID-19)"))

Fig4ruca_cv19_barriers <- ggplot (CV19barriers_ruca, aes (x=RUCA, y=Percent, fill= ff)) + 
  geom_bar (position = "stack", stat = "identity", width = 0.8) +
  facet_wrap (~fct_relevel (Rank, "Rank2", "Rank5", "Rank3", "Rank4", "Rank1"), nrow=5) +
  

  coord_flip() +

          scale_y_continuous(name="Percentage of Providers", labels = function (y) paste0(y, "%")) +
    theme_minimal () +
   theme(axis.text.x = element_text(face="bold", colour = "black",
                           size=14, angle=0),
          axis.text.y = element_text(face="bold", colour = "black",
                           size=13, angle=0)) +
  #geom_text (aes (label=paste0 (round (Percent, 1), "%")), size = 5, hjust = 0.8) +
  theme(legend.title=element_blank(),
        legend.text = element_text (size = 12, face="bold"),
        axis.title.y = element_blank (),
        axis.title.x = element_text (size = 16, face = "bold")) +
  theme (legend.position = "bottom") +
  theme (axis.title.x = element_text (margin = margin (t=20)))+
  theme (panel.grid.major = element_line (size = 0.5, linetype = "solid", color="grey"))+
  theme (panel.spacing = unit (2,"lines")) +
  theme (strip.text.x = element_blank () ) +
 scale_fill_brewer(palette = "Paired") +
   guides (fill = guide_legend (reverse = TRUE))


pdf ("Fig4ruca_cv19_barriers.pdf")
Fig4ruca_cv19_barriers
dev.off ()
```

